---
title: "P8105 Homework 1"
author: "Zheming Bao"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 7, fig.height = 5)
library(ggplot2)
```

## Problem 1

```{r load_penguins}
data("penguins", package = "palmerpenguins")
```

The `penguins` dataset has `r nrow(penguins)` rows (one penguin per row)
and `r ncol(penguins)` columns. It covers Antarctic penguins of three species:
`r paste(levels(penguins$species), collapse = ", ")`, and the islands are
`r paste(levels(penguins$island), collapse = ", ")`. Measurements include
bill length and depth (mm), flipper length (mm), and body mass (g).
It also records sex (female or male, with missing values)
and year (`r min(penguins$year)`–`r max(penguins$year)`).

The mean flipper length is
`r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2)` mm.
This excludes `r sum(is.na(penguins$flipper_length_mm))` missing values.

```{r penguin_scatterplot}
penguin_plot = ggplot(
  penguins,
  aes(x = bill_length_mm, y = flipper_length_mm, color = species)
) +
  geom_point(alpha = 0.75, na.rm = TRUE) +
  labs(
    title = "Bill length and flipper length by penguin species",
    x = "Bill length (mm)",
    y = "Flipper length (mm)",
    color = "Species"
  ) +
  theme_minimal()

penguin_plot

ggsave(
  filename = "penguin_scatterplot.png",
  plot = penguin_plot,
  width = 7,
  height = 5,
  units = "in",
  dpi = 300
)
```

Each point is a penguin. Species determines discrete colors, while numeric
measurements use continuous axes. Gentoo penguins generally have longer
flippers. Missing coordinates are omitted; the plot is saved as
`penguin_scatterplot.png`.

## Problem 2

```{r create_dataframe}
set.seed(2367)

sample_data = data.frame(
  normal_sample = rnorm(10),
  character_value = letters[1:10],
  factor_value = factor(
    rep(c("low", "medium", "high"), length.out = 10),
    levels = c("low", "medium", "high")
  )
)

sample_data$positive = sample_data$normal_sample > 0
sample_data = sample_data[
  c("normal_sample", "positive", "character_value", "factor_value")
]

sample_data
```

The ten rows contain standard Normal draws, an indicator for values above
zero, letters, and a three-level factor. The seed makes the sample reproducible.

```{r variable_means}
mean(sample_data$normal_sample)
mean(sample_data$positive)
mean(sample_data$character_value)
mean(sample_data$factor_value)
```

The numeric mean is `r round(mean(sample_data$normal_sample), 3)`.
The logical mean is `r mean(sample_data$positive)`, the proportion above zero.
Character and factor means return `NA` with warnings because these types
are not numeric or logical.

```{r numeric_coercion, results='hide', warning=FALSE}
as.numeric(sample_data$positive)
as.numeric(sample_data$character_value)
as.numeric(sample_data$factor_value)
```

Logical values become 1 (`TRUE`) or 0 (`FALSE`), explaining their proportion
mean. Letters become `NA` with a warning because they are not numeric strings.
Factors become level codes (`low` = 1, `medium` = 2, `high` = 3). These codes
label categories, so their mean is not a meaningful category average;
`mean()` on the original factor returns `NA`.
