---
title: "P8105 Homework 1"
author: "Zheming Bao"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 7, fig.height = 5)
library(ggplot2)
```

## Problem 1

```{r load_penguins}
data("penguins", package = "palmerpenguins")
```

The `penguins` dataset contains measurements of penguins from the Palmer
Archipelago in Antarctica. It has `r nrow(penguins)` rows and
`r ncol(penguins)` columns, with one penguin per row. The species are
`r paste(levels(penguins$species), collapse = ", ")`, and the islands are
`r paste(levels(penguins$island), collapse = ", ")`. Measurements include
bill length and depth (mm), flipper length (mm), and body mass (g).
The dataset also records sex (female or male, with some missing values)
and year (`r min(penguins$year)`–`r max(penguins$year)`).

The mean flipper length is
`r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2)` mm.
This calculation excludes `r sum(is.na(penguins$flipper_length_mm))`
missing flipper-length measurements, using the remaining
`r sum(!is.na(penguins$flipper_length_mm))` observations.

```{r penguin_scatterplot}
penguin_plot = ggplot(
  penguins,
  aes(x = bill_length_mm, y = flipper_length_mm, color = species)
) +
  geom_point(alpha = 0.75, na.rm = TRUE) +
  labs(
    title = "Bill length and flipper length by penguin species",
    x = "Bill length (mm)",
    y = "Flipper length (mm)",
    color = "Species"
  ) +
  theme_minimal()

penguin_plot

ggsave(
  filename = "penguin_scatterplot.png",
  plot = penguin_plot,
  width = 7,
  height = 5,
  units = "in",
  dpi = 300
)
```

Each point represents a penguin, with species mapped to color. Species is
a factor, so the colors form discrete groups. Bill and flipper lengths are
numeric, so both axes are continuous. Gentoo penguins generally have longer
flippers than the other two species. The plot omits rows with missing
coordinates and is saved as `penguin_scatterplot.png` in the project directory.
