---
title: "P8105 Homework 1"
author: "Zheming Bao"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 7, fig.height = 5)
library(ggplot2)
```

## Problem 1

```{r load_penguins}
data("penguins", package = "palmerpenguins")
```

The `penguins` dataset contains measurements of penguins from the Palmer
Archipelago in Antarctica. It has `r nrow(penguins)` rows and
`r ncol(penguins)` columns, with one penguin per row. The species are
`r paste(levels(penguins$species), collapse = ", ")`, and the islands are
`r paste(levels(penguins$island), collapse = ", ")`. Measurements include
bill length and depth (mm), flipper length (mm), and body mass (g).
The dataset also records sex (female or male, with some missing values)
and year (`r min(penguins$year)`–`r max(penguins$year)`).

The mean flipper length is
`r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2)` mm.
This calculation excludes `r sum(is.na(penguins$flipper_length_mm))`
missing flipper-length measurements, using the remaining
`r sum(!is.na(penguins$flipper_length_mm))` observations.

```{r penguin_scatterplot}
penguin_plot = ggplot(
  penguins,
  aes(x = bill_length_mm, y = flipper_length_mm, color = species)
) +
  geom_point(alpha = 0.75, na.rm = TRUE) +
  labs(
    title = "Bill length and flipper length by penguin species",
    x = "Bill length (mm)",
    y = "Flipper length (mm)",
    color = "Species"
  ) +
  theme_minimal()

penguin_plot

ggsave(
  filename = "penguin_scatterplot.png",
  plot = penguin_plot,
  width = 7,
  height = 5,
  units = "in",
  dpi = 300
)
```

Each point represents a penguin, with species mapped to color. Species is
a factor, so the colors form discrete groups. Bill and flipper lengths are
numeric, so both axes are continuous. Gentoo penguins generally have longer
flippers than the other two species. The plot omits rows with missing
coordinates and is saved as `penguin_scatterplot.png` in the project directory.

## Problem 2

```{r create_dataframe}
set.seed(2367)

sample_data = data.frame(
  normal_sample = rnorm(10),
  character_value = letters[1:10],
  factor_value = factor(
    rep(c("low", "medium", "high"), length.out = 10),
    levels = c("low", "medium", "high")
  )
)

sample_data$positive = sample_data$normal_sample > 0
sample_data = sample_data[
  c("normal_sample", "positive", "character_value", "factor_value")
]

sample_data
```

The data frame has ten rows and four columns: a sample from a standard
Normal distribution (mean 0 and standard deviation 1), a logical indicator
of whether each sampled value is greater than zero, ten character values,
and a factor with three levels. Setting a seed makes the sample reproducible.

```{r variable_means}
mean(sample_data$normal_sample)
mean(sample_data$positive)
mean(sample_data$character_value)
mean(sample_data$factor_value)
```

The numeric mean is
`r round(mean(sample_data$normal_sample), 3)`.
The logical mean is `r mean(sample_data$positive)`, the proportion of sampled
values greater than zero: R treats `TRUE` as 1 and `FALSE` as 0 when taking
this mean. The character and factor means both return `NA` with warnings
because `mean()` does not accept these variable types as numeric measurements.

```{r numeric_coercion, results='hide', warning=FALSE}
as.numeric(sample_data$positive)
as.numeric(sample_data$character_value)
as.numeric(sample_data$factor_value)
```

The logical values convert to 1 (`TRUE`) and 0 (`FALSE`), explaining why
their mean is a proportion. The character values here are letters rather
than numeric strings, so they convert to `NA` and normally produce a
warning. Numeric strings such as `"12"` could be converted, but these letters
cannot. Output and warnings are hidden in this chunk as requested.

The factor converts to its internal level codes: `low` becomes 1, `medium`
becomes 2, and `high` becomes 3, according to the explicitly specified level
order. These codes identify categories; they do not establish numerical
distances between the categories. Although the converted codes have a
computable mean, it is not a meaningful average of the original categories.
Thus, having internal integer codes does not make a factor a numeric
measurement, and taking the mean of the original factor returns `NA`.
